---
title: "BIO2090 Exam data"
author: "Nicholas Harmer"
date: "4 August 2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a script to generate unique data sets for the end of module examination.It should provide each student with unique data, and provide markers with an answer sheet.

Three packages are required. ggplot2 (https://ggplot2.tidyverse.org/authors.html) is required to prepare images for the marker. officer (https://davidgohel.github.io/officer/index.html) provides the ability to write and amend files in Word format. magrittr (https://magrittr.tidyverse.org/) makes the use of officer more straightforward.

```{r}
# Load packages
library(ggplot2)
library(officer)
library(magrittr)
```

The working directory and list of students are set up (using dummy numbers in this example). The main parameter that might be changed is the error ranges that are used. The two values are estimated errors in pipetting and in data reading respectively. The values chosen robustly produce data that is readily interpreted. This reflects data produced in the past from the top quartile of students. Higher values would produce more challenging data (perhaps suitable for formative assessment); lower values would produce data that were more "perfect" if this were required.
```{r}
setwd("C:/Users/njh209/OneDrive - University of Exeter/Nic's documents/R/resources")

# Load student numbers
Students = c("C12345", "C23456")

# Set errors
error <- c(15, 7)
```

Set up a loop to run the script for all students. For a test, change the student numbers to a list of one or use the second (commented out) line. The loop is not set up in its entirety here to allow comment on the remainder of the loop. To run, move the end of the loop to the end of the script.
```{r}
for (j in 1:length(Students)) {
}
```

The Bradford data is then generated. This will be required for later calculations so it is important to generate this first. The data are set up as a data frame: this is most convenient for later providing to students and markers in a table. The first three elements of the data frame define the experiment. The second three elements are data to be generated and so are set to dummy values.
Two random elements are then generated, for the background reading and the gradient of the data generated. In the laboratory these show batch to batch variation from the preparation of Bradford reagent and the variability partially reflects this.
```{r}
  # Generate Bradford data
  Bradford <- data.frame(Tube = c(1:6),
                         Volume_of_BSA_solution_uL = c(0,4,8,12,16,20),
                         Volume_of_water_uL = c(1000,996,992,988,984,980),
                         Absorbance_595nm = c(0),
                         concentration = c(0),
                         Corrected_Absorbance_595nm = c(0)) 
  grad <- 0.0025+(sample(0:100,1)/100000)
  back <- 0.9+(sample(0:200,1)/1000)
```

The Bradford data are then generated by a loop. The observed absorbance is generated from the sample concentrations, applying the errors defined earlier. These will be provided to students. The correct answers for sample concentration and background corrected readings are generated here to be provided to markers.
```{r}
  for (i in 1:6) {
    Bradford[i,4] <- round(((100-error[2]+((sample(0:(error[2]*200),1))/100))/100)*(back+(Bradford[i,2]*10
                                    *grad*((100-error[1]+((sample(0:(error[1]*200),1))/100))/100))), digits = 3)
    Bradford[i,5] <- Bradford[i,2]*10
    Bradford[i,6] <- Bradford[i,4]-Bradford[1,4]
  }  
```

A linear regression of the correct answers is then performed. This will be used to calculate correct answers for later questions. The lm function performs linear regression on the data in the "Bradford" data frame and outputs the answers to a list. For convenience I copied the relevant outputs to individual variables to help me recall what I am doing. 
```{r}
  # Calculate linear regression of Bradford
  Brlr <- lm(Bradford[,6]~Bradford[,5])
  slope <- as.numeric(Brlr$coefficients[2])
  yint <- as.numeric(Brlr$coefficients[1])
```

The next part of the script generates data for the "unknown" samples that students would prepare in the practical class. Again, this is set up as a data frame to facilitate provision of data to students and markers. Three values are defined by the experiment; the absorbance for each sample is randomly determined within ranges that we have previously observed; and dummy values are assigned where calculations will be necessary.
```{r}
  # Generate table of results for unknowns
  Fractions <- data.frame(Fraction = c("Homogenate, H", "Cytosol, C", "Mitochondria, M", "Nuclear, N",
                                       "30% ammonium sulfate", "50% ammonium sulfate", "80% ammonium sulfate"),
                          Volume_of_sample_uL = c(5),
                          Volume_of_water_uL = c(996),
                          Absorbance_595nm = c((back+0.35+sample(0:200,1)/1000), 
                                               (back+0.5+sample(0:300,1)/1000),
                                               (back+0.4+sample(0:400,1)/1000),
                                               (back+0.25+sample(0:300,1)/1000),
                                               (back+0.3+sample(0:50,1)/1000),
                                               (back+0.5+sample(0:200,1)/1000),
                                               (back+0.3+sample(0:100,1)/1000)),
                          Corrected_Absorbance = c(0),
                          x_value = c(0),
                          concentration = c(0))
```

A loop is then run to perform calculations to generate the remaining data. The first calculation provides the background corrected value; the second calculates the correct value from the linear regression (which is provided to the marker so that students who make an error in the following step can be rewarded for correct caluclation); and the last step calculates the final answer. These will be provided to the marker. The "signif" function is used to provide the data to an appropriate number of significant figures. 
```{r}
  for (i in 1:7) {
    Fractions[i,5] <- Fractions[i,4]-Bradford[1,4]
    Fractions[i,6] <- signif((Fractions[i,5]-yint)/slope,4)
    Fractions[i,7] <- signif(Fractions[i,6]/5,3)
  }
```

This process is repeated for the PGK assay data. Again the values provided to the students are chosen from a range that reflects past student data. A loop is then used to perform the calculations to convert these data as required by the exam questions.
```{r}
  # Generate PGK data
  PGK <- data.frame(Fraction = c("Homogenate, H", "Cytosol, C", "Mitochondria, M", "Nuclear, N",
                                 "30% ammonium sulfate", "50% ammonium sulfate", "80% ammonium sulfate"),
                    Gradient_of_slope_AU_per_min = c((-0.03-sample(0:200,1)/1000), 
                                         (-0.03-sample(0:200,1)/1000),
                                         (sample(0:100,1)/-1000),
                                         (-0.1-sample(0:100,1)/1000),
                                         (-0.03-sample(0:20,1)/1000),
                                         (-0.03-sample(0:30,1)/1000),
                                         (-0.15-sample(0:200,1)/1000)),
                    Product_concentration = c(0),
                    Activity = c(0),
                    Mass_protein = c(0),
                    Specific_activity = c(0))
  
  for (i in 1:7) {
    PGK[i,3] <- signif(PGK[i,2]*1000000/-6220,5)
    PGK[i,4] <- PGK[i,3]/1000
    PGK[i,5] <- Fractions[i,7]*0.005
    PGK[i,6] <- signif(PGK[i,4]/PGK[i,5],3)
  }
```

The student output file is generated. Firstly, a new document is generated. The following line invokes the magrittr utility to make adding further text convenient. A title is given to the file, and the data are then written to the file. These are the student's number; the Bradford standard data (only columns 1-4); the sample Bradford data (again, only columns 1-4); and the sample PGK assay data (only columns 1-2).
The file is then written out into a "Students" folder with the student number as the filename.
```{r}
  # Generate student files

  readout <- read_docx()
  readout <- readout %>%
    body_add_par("BIO2090 2020/21 January Examination", style = 'heading 1') %>%
    body_add_par("") %>%
    body_add_par(paste0("Individual data for student: ", Students[j]," for Question 1")) %>%
    body_add_par("") %>%
    body_add_par("Bradford standard curve data:", style = "Normal") %>%
    body_add_par("") %>% 
    body_add_table(value = Bradford[,1:4], style="table_template") %>%
    body_add_par("") %>%
    body_add_par("") %>%
    body_add_par("Sample Bradford data:") %>%
    body_add_par("") %>%
    body_add_table(value = Fractions[,1:4], style="table_template") %>%
    body_add_par("") %>%
    body_add_par("") %>%
    body_add_par("PGK assay data:") %>%
    body_add_par("") %>%
    body_add_table(value = PGK[,1:2], style="table_template") %>%
    

  print(readout, target = paste0("Student_files/", Students[j], ".docx"))
```

The marker file will include a figure of the Bradford assay standard. The ggplot package produces beautiful figures and is very well supported with documentation. This is used to generate the plot. The linear regression algorithm used earlier is again used to determine the line of best fit, and in this case provides also a confidence interval. 
```{r}
  # Prepare figures
  se <- ggplot(data = Bradford, aes(x = concentration, y=Corrected_Absorbance_595nm)) + 
    stat_smooth(method = "lm",  formula = y~x, level = 0.75)
  se <- se + geom_point(mapping = aes(size = 2)) + labs(y="Absorbance / AU", x="concentration / ug/mL")
```

Finally, the marker files are prepared. A new document is generated. This gives the student number; the full data table for the Bradford standards with the correct calculations included; the gradient and intercept for the line of best fit; the figure for the Bradford standards; the full table for the sample Bradford data, including answers and working; and the full table for the PGK assay, including answers and working. These are written to a "Lecturer" folder with the student number again as the filename.
```{r}
  # Generate lecturer files
  
  readout <- read_docx()
  readout <- readout %>%
    body_add_par("BIO2090 2020/21 Exam Dataset", style = 'heading 1') %>%
    body_add_par("") %>%
    body_add_par(paste0("Data and answers for student: ", Students[j])) %>%
    body_add_par("") %>%
    body_add_par("") %>%
    body_add_par("Bradford standard curve data:", style = "Normal") %>%
    body_add_par("") %>% 
    body_add_table(value = Bradford, style="table_template") %>%
    body_add_par("") %>%
    body_add_par("") %>%
    body_add_par(paste0("Bradford graph: formula is y = ", signif(slope,3),"x + ",signif(yint,3))) %>%
    body_add_par("") %>%
    body_add_gg(value = se, style = "centered") %>%
    body_add_par("") %>%
    body_add_par("") %>%
    body_add_par("Sample Bradford data:") %>%
    body_add_par("") %>%
    body_add_table(value = Fractions, style="table_template") %>%
    body_add_par("") %>%
    body_add_par("") %>%
    body_add_par("PGK assay data:") %>%
    body_add_par("") %>%
    body_add_table(value = PGK, style="table_template") %>%
    

  print(readout, target = paste0("Lecturer_files/", Students[j], "_answers.docx"))
```
Finally, the loop is closed so that all of these actions are run for every student.
